### 2015-06-15 writeDown/pub.epj/comment.klein.reply/main.tex
add reply to Klein

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal Special Topics
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[epjST]{svjour}
%
\usepackage{graphics}
\usepackage{color, colortbl}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{tabularx}

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{red} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dd}[1]{\textsf{#1}}
\newcommand{\fwd}[0]{\textrm{fw}}
\newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mh}[0]{\mathcal {H}}
\newcommand{\dist}[0]{\textrm {dist}}
\newcommand{\AT}[0]{\textrm{AT}}
\newcommand{\HY}[0]{\textrm{HY}}
\newcommand{\CG}[0]{\textrm{CG}}
\newcommand{\EX}[0]{\textrm{EX}}
\newcommand{\moleidxone}[0]{i}
\newcommand{\moleidxtwo}[0]{j}
\newcommand{\atomidxone}[0]{\alpha}
\newcommand{\atomidxtwo}[0]{\beta}
\newcommand{\thf}{{\textrm{th}}}
\newcommand{\rdf}{{\textrm{rdf}}}
\newcommand{\rep}{{\textrm{rep}}}
\newcommand{\dof}{{\textrm{DOF}}}
\newcommand{\exc}{{\textrm{extra}}}
\newcommand{\equi}{{\textrm{eq}}}

\newcommand{\confa}[0]{{\alpha_{\textrm{R}}}}
\newcommand{\confb}[0]{{\textrm{C}7_{\textrm{eq}}}}
\newcommand{\confc}[0]{{\alpha_{\textrm{L}}}}
\newcommand{\confd}[0]{{\textrm{C}7_{\textrm{ax}}}}

% \usepackage{amsfonts}
% \newcommand{\tickYes}{\checkmark}
% \usepackage{pifont}
% \newcommand{\tickNo}{\hspace{1pt}\ding{55}}
% \definecolor{MyGray}{gray}{0.85}

%
\begin{document}
%
\title{Reply to comment by R.~Klein on ``Adaptive Resolution Simulation in Equilibrium and Beyond''}
\author{Han Wang\inst{1}\fnmsep\thanks{\email{wang\_han@iapcm.ac.cn}}, Animesh Agarwal\inst{2}}
%
\institute{CAEP Software Center for High Performance Numerical Simulation, Beijing, China
  \and Institut f\"ur Mathematik, Freie Universit\"at Berlin, Berlin, Germany}
%
% \abstract{
%   We reply comments~\cite{klein2015comment}
%   on our recent paper~\cite{wang2015adaptive}.
% } %end of abstract
%
\maketitle
%

We would like to thank R.~Klein who provided valuable comments~\cite{klein2015comment}
on our recent paper~\cite{wang2015adaptive}.
The questions raised therein will be replied in the following.

\paragraph{Particles in the hybrid region}
In the AdResS Scheme, the minimum thickness of the hybrid region
should be $r_c$ (cut-off radius) if the traditional weighting function
(i.e.~Eq.1 in \cite{wang2015adaptive}) is used, while it should be at least
$2r_c$ if a buffer region  is added to the hybrid
region (i.e.~Eq.8 in the same reference).
The average molecular spacing of liquid water under ambient
condition is about 0.3~nm. Therefore, if one take 1.2~nm as the cut-off
radius, the thickness is 4 and 8 times the
the average molecular spacing for traditional and buffered hybrid region, respectively.

We assume the interactions between particles are pairwise, and denote
the energy contribution of one particle pair by $u(r)$, with $r$ being the distance
between the paticles. Cutting-off this interaction means that it is assumed to vanish
if the distance between the paticles is larger than the cut-off radius $r_c$.
``The correlation between the atomistic and the coarse-grained
regions is negligible'' means that the
degrees of freedomes (positions and velocities) of two paticles
in the atomistic and the coarse-grained regions, respectively, are almost un-correlated.

It is true that by intrducing the adaptive resolution, the system is not homogeneous
and isotropic anymore. Therefore, strictly speaking the definition
of RDF and 3-body correlation is problematic. However, if the configuration
of the coarse-grained and the
hybrid regions are designed with high accuracy, one can approximately
treat the system as homogeneous and isotropic, and the RDF and 3-body correlation
are well-defined for (at least) the center-of-mass configuration of water system.
On the other hand, if the coarse-grained or hybrid
configuration deviates from the full-atomistic reference,
one can anyway computes the RDF,
which is the averged structure around a particle, and contains information on
the difference to the
full-atomistic reference (probably with positive-negative cancellation).
In practice,
the deviating RDFs is to improve the accuracy of AdResS, and a
satisfactory accuracy of RDF and 3-body correlation is achieved by
using a fully converged RDF correction force in
AdResS~\cite{wang2012adaptive,wang2013grand}.

\paragraph{The thermodynamic force}
The thermodynamic force in the atomistic and the coarse-grained
regions are force to vanish, that means if a zero thermodynamic force
is used initially, only the value in the hybrid region is updated
iterativly. In order to calculate the density,
take the pure TIP3P water system in Ref~\cite{wang2015adaptive}
for example, we did the following:
\begin{enumerate}
\item The hybrid region was divided into bins of width 0.01~nm in the
  direction of resolution changing.
\item The number of molecules (center-of-mass position) falling in
  each bin was counted and averged to calculate a density. Each density data point
  has a statistical uncertainty of roughly 8\%.
  The productive trajectory for the computation was 200~ps long
  and the molecular positions were saved every 0.2~ps.
\item The resulting density profile was still oscillatory,
  so we smooth it by fitting the density
  to a  spline with a grid spacing of 0.4~nm.
  The spline was calculated by software VOTCA~\cite{ruehle2009versatile}.
\item This smoothed density was used
  by the iterative algorithm (Eq.~21 in Ref.~\cite{wang2015adaptive}),
  % that computes the
  and the thermodynamic force converges in 10 iterations.
\end{enumerate}

\paragraph{Particular properties of the SPC/E water system}
Actually we used the AdResS simulation to compute the chemical
potential for various liquids and mixtures in
Ref.~\cite{agarwal2014chemical}.  In the simulations, only the
thermodynamic force was calculated and applied to correct the density
in the hybrid region.  The chemical potential were computed with
satisfactory accuracy.  Moreover, although no RDF correction was used
and the system was simulated under extreme technical conditions (very
small atomistic and coarse-grained regions and a relatively large
hybrid region), the RDFs were be reproduced with satisfactory
agreement to full-atomistic reference (see e.g.~Fig.3 of Ref.~\cite{agarwal2014chemical}).

\paragraph{Structure of the hybrid zone}
The major difference between a phase coexisting system and an AdResS
system is that the former satisfies both the momentum and energy
conservations, while the later can satisfy only one of the two.
It would be very intersting to establish the analogue between
the phase coexisting and the AdResS system, however one must pay attention
to this difference that may play a significant role in the final conclusions.

\bibliography{ref}{}
\bibliographystyle{unsrt}

% \begin{thebibliography}{}
% % and use \bibitem to create references.
% \bibitem{RefJ}
% % Format for Journal Reference
% Author, Journal \textbf{Volume}, (year) page numbers
% % Format for books
% \bibitem{RefB}
% Author, \textit{Book title} (Publisher, place year) page numbers
% % etc
% \end{thebibliography}

\end{document}

% end of file template.tex